---
title: "Analysis of fairness based on talent and value in the hiring and remuneration processes"
subtitle: "A special focus on gender parity"
author: "Report prepared for Black Saber Software by SFP"
date: '2021-04-21'
output:
  pdf_document:
    template: report.tex
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
lang: en
titlepage: yes
titlepage-color: 6C3082
titlepage-text-color: FFFFFF
titlepage-rule-color: FFFFFF
titlepage-rule-height: 2
---

```{r, message = FALSE, echo=FALSE}
library(tidyverse)
library(lme4)
library(gtsummary)
library(kableExtra)
library("FrF2")
# this should supress all code and messages
knitr::opts_chunk$set(echo = FALSE)
```


\newpage
# Executive summary

_Guidelines for the executive summary:_

**Background & Aim** :

Concerns about potential bias in the hiring and remuneration processes were consistently raised internally in Black Saber Software. In order to investigate the true fact for the issues, we SFP were consent to help figure it out. This report was commissioned to examine if potential issues especially gender parity existed in the hiring, promotion and salary processes based on the hiring data and data about promotion and wages for current employees provided by Black Saber Software . Besides, we were interested in the fact how the new trialled AI service performanced in the hiring process. 

**Key results** \
final result vs. productivity \
final hiring result vs. gender \
AI result good? \
promotion vs. productivity \
promotion vs. gender \
salary vs. productivity \
salary vs. gender

**Limitations** :

We were only offered with data for current employees so it was hard to know reasons past employees left. If additional data related to past employees can be provided, we can have a more deep and comprehensive sight into potential biases;
The lack of ethnicity/race data also results in a limitation since race might be a potential exist which might influence the results of hiring pipelines and promotion and amount of wages;
No information about the determination methods for hiring results, promotion and amount of salary was provided which limited the accuracy of models we fit for these processes.

\newpage
Key results are summarized in the following tables or graphs.


\newpage
# Technical report


## Introduction

Recently, several complaints were received related to potential bias in the hiring and remuneration processes. Hence, BSS (Black Saber Software) required our team to take a look and give out a related report for the Board of Directors. The purpose of this study is to investigate if BSS’s hiring, promotion, and salary were processing fairly based on talent and value to the company. We were especially interested in exploring potential gender bias in the hiring and remuneration processes. Besides, we also investigated the performance of the new AI service in the hiring pipeline. Data was provided by the data team from BSS including hiring data for their new graduate program and data for their current employees. 

In this report, we will specify different explanatory and response variables for. Then, we will fit several generalized linear models and a linear mixed effects model to analyze the results. At the end, we will point out the same limitations of our model. The whole report will be run on R.studio and knit into pdf.


## Research questions
* Does the new AI service in the hiring pipeline work and is reliable for future use?
* In particular, is gender a potential bias during the promotion and salary processes?
* If the promotion, and salary process are all fair to all employees and based on their ability and value to BSS?

\newpage

## Models and Results
### Hiring Process for applicants
**Relationships between whether a candidate was hired or not and the variables during the entire hiring process**

We will consider a data set named all_phase combining all necessary data in the hiring process which include four datasets : phase1-new-grad-applicants-2020.csv, phase2-new-grad-applicants-2020.csv, phase3-new-grad-applicants-2020.csv and final-hires-newgrad_2020.csv. Each line of all_phase contains all aggregated information of an applicant in the hiring pipelines. In this analysis for hiring process, the response variable is the 
Examination of data for hiring process reveals the following key variables in all_phase:

applicant_id = Unique identifier assigned to candidates in Phase 1 \
team_applied_for = Data or Software \
gpa = 0.0 to 4.0 \
extracurriculars = Indicates the quality of extracurricular involvement with scores of 0,1,2 \
work_experience = Description of candidates’ previous work experience with scores of 0,1,2 \
technical_skills = Score from 0 to 100 on a timed technical task, AI autograded \
writing_skills = Score from 0 to 100 on a timed writing task, AI autograded \
speaking_skills = A rating of speaking skills based on pre-recorded video, AI autograded \
leadership_presence = A rating of leadership presence based on pre-recorded video, AI autograded \
final_result = Indicate at which phase applicant was rejected or hired finally \
result = Show the hiring result for each applicant in each hiring pipeline 



```{r, warning=FALSE, message=F}
# read in the data
black_saber_current_employees <- read_csv("data/black-saber-current-employees.csv")
phase1 <- read_csv("data/phase1-new-grad-applicants-2020.csv")
phase2 <- read_csv("data/phase2-new-grad-applicants-2020.csv")
phase3 <- read_csv("data/phase3-new-grad-applicants-2020.csv")
final <- read_csv("data/final-hires-newgrad_2020.csv")
#join three datasets into one.
phase1_2 <- full_join(phase1,phase2)
all_phase <- left_join(phase1_2, phase3, by = "applicant_id")
final_applicant <- left_join(final, all_phase, by = "applicant_id") 
all_phase$final_result <- ifelse(all_phase$applicant_id %in% final_applicant$applicant_id, "Hired",          
                           ifelse(all_phase$applicant_id %in% phase3$applicant_id,"Refused at phase 3",  
                           ifelse(all_phase$applicant_id %in% phase2$applicant_id,"Refused at phase 2","Refused at phase 1")))
all_phase$final_result <- factor(all_phase$final_result,level=c("Refused at phase 1","Refused at phase 2","Refused at phase 3","Hired"))
# create new variables to show the hiring results for each candidates in each phrase
phase1$result <- ifelse(phase1$applicant_id %in% phase2$applicant_id,"proceed","refused")
phase2$result <- ifelse(phase2$applicant_id %in% phase3$applicant_id,"proceed","refused")
phase3$result <- ifelse(phase3$applicant_id %in% final$applicant_id,"proceed","refused")
all_phase$result <- ifelse(all_phase$applicant_id %in% final$applicant_id,"Hired","no")
# Create model 1 to see if several response variables are related to the final result.
```


```{r, warning=F}
model_1 <- glm(as.factor(result) ~ gpa+cv+cover_letter+extracurriculars+work_experience,
               data = all_phase, family = binomial("logit"))
tbl_regression(model_1)
#Create model 2 to see AI score preference
```

**Is the AI used during the hiring process reliable.**
```{r}
model_2 <- glm(as.factor(result) ~ gpa+extracurriculars+work_experience+leadership_presence+
                 speaking_skills+technical_skills+writing_skills, 
               data = phase2, family = binomial("logit"))
tbl_regression(model_2)
```

\newpage 

### Promotion and Salary Processes for current employees

In the current employee data set, we have selected gender, team, leadership_for_level, productivity, and salary. Based on the label dictionary, obtained along with the data, each variables represent the following:

employee_id = Unique 5-digit employee identifier \
gender =  Gender of employee: man, woman, prefer not to say \
team = one of the 8 teams which each employee works for \
leadership_for_level =  if the employee is appropriate for level, need improvements or exceeds expectations \
productivity = work output with respect to job description rated on a 0 - 100 scale \
salary = salary at the given financial quarter 

```{r}
## gender, promotion, and salary
# clean the salary data for further analysis
black_saber_current_employees$salary<-gsub("\\$","",black_saber_current_employees$salary)
black_saber_current_employees$salary<-gsub("\\,","",black_saber_current_employees$salary)
black_saber_current_employees$salary<-as.numeric(black_saber_current_employees$salary)
# change the data types
black_saber_current_employees$employee_id<-as.character(black_saber_current_employees$employee_id)
black_saber_current_employees$gender<-as.factor(black_saber_current_employees$gender)
black_saber_current_employees$team<-as.factor(black_saber_current_employees$team)
black_saber_current_employees$role_seniority<-as.factor(black_saber_current_employees$role_seniority)
black_saber_current_employees$leadership_for_level<-as.factor(black_saber_current_employees$leadership_for_level)
```

```{r}
# check general relationship between productivity and salary for each team
black_saber_current_employees %>% 
 ggplot(aes(x= productivity,y=salary,color=gender))+
  geom_smooth(method = "lm",
              se = FALSE,
              size = 1.5) +
  facet_wrap(~team) +
  theme_minimal()+
  labs(title="Productivity by salary",
       x="Productivity",
       y="Salary",
       caption = str_c("Prepared by SFP\n",Sys.Date()))
ggsave("images/productivity_salary.png",width = 7, height = 4)
```

```{r}
#models to check if salary is related to other variables
mod_1 <-lmer(salary~gender+team+(1|employee_id),
             data=black_saber_current_employees)
summary(mod_1)

#tab_model(mod_1)
tbl_regression(mod_1)

```


```{r}
mod_2 <-lmer(salary~gender+team+role_seniority+productivity+leadership_for_level+(1|employee_id:team)+(1|employee_id),data=black_saber_current_employees)
lmtest::lrtest(mod_1,mod_2)
lrt_table_data <- data.frame(unclass(lmtest::lrtest(mod_1,mod_2)), check.names = FALSE, stringsAsFactors = FALSE)
lrt_table_data['LogLik'] <- round(lrt_table_data['LogLik'], 0)
lrt_table_data['Chisq'] <- round(lrt_table_data['Chisq'],0)
lrt_table_data <- sapply(lrt_table_data, as.character)
lrt_table_data[is.na(lrt_table_data)] <- " "

kable( lrt_table_data, align = "lllll", booktabs = T) %>%
  kable_styling(full_width = T)
```

```{r}
#check general relationship between productivity and leadership_for_level for each team
black_saber_current_employees %>% 
 ggplot(aes(x= leadership_for_level,fill=gender))+
  geom_bar(position="fill")+
  theme_minimal()+
  facet_wrap(~team)+
  labs(title="Gender distribution on leadership for level",
       x="Leadership for level",
       y="Percentage",
       caption = str_c("Prepared by SFP\n",Sys.Date())) +
  scale_x_discrete(guide = guide_axis(n.dodge = 3))
  ggsave("images/productivity_leadership.png",width = 7, height = 4)
```


## Discussion

First, we explored the effects the various variables during the hiring phases had on the hiring chances. Fitting a model to model the applicant's hiring result (no if not hired, yes if hired) against the applicant's GPA,Curriculum Vitae, cover letter, extracurriculars, and work_experience, it was found that only the GPA and extracurriculars had a significant effect on the being hired, with p-values of 0.009 and 0.042 respectively.

Second, we explored whether the AI used during the first 2 phases of the hiring process was reliable, and if there were any biases.

## Strengths and limitations

One limitation is that we were only provided data on current employees. The data of previous past employees is missing. Regarding the promotion and salary processes, past employees’ information could give us some insights to why they left the company. This could give valuable information such as whether they were treated evenly. If we can get the past employees’ data, we can analyze whether there exist biases during the remuneration process more comprehensively.


Another limitation results from the lack of ethnicity/race data for both the hiring and remuneration processes which could also be a potential bias that we need to check. We were told that the race data was not collected by the People and Talent team but they were considering it for EDI initiatives. The race might also be a factor that would affect the hiring result of applicants and also the promotion and wages processes of employees. 


Finally, a limitation may be introduced as a result of implicit evaluation criterions for each hiring phase, especially AI-autograded phase  and remuneration process. If we know how applicants can proceed to the next round in the hiring process, it would be easier to identify what effects are correlated and fit better models. For the remuneration process, if we can know what kinds of employees can be promoted or have higher salary, we will be able to explore more about potential biases.



\newpage
# Consultant information
## Consultant profiles

**Lucas Xian**. Lucas is a senior consultant with SFP. He specializes in data analysis and visualization. Lucas earned his Honours Bachelor of Science, Specialist in Statistics, from the University of Toronto in 2023.

**Dandan Zhang**. Dandan is a junior consultant with SFP. She earned her Bachelor of Science, majoring in Mathematics and Statistics, from the University of Toronto in 2021.She specializes in data visualization. 

**Jong-Hoon Kim** John is a senior consultant with SFP. He earned his Bachelor of Science in statistics and economics, from the University of Toronto in 2021. He specialises in data analysis and visualisation. 

**Zikun Lei** Zikun Lei is a senior consultant with SFP. He earned his Bachelor of Science in statistics and economics, from the University of Toronto in 2021. He specialises in data analysis and visualisation. 


## Code of ethical conduct

SFP is a statistical consulting company. Our mission is to help our clients solve their problems and give them professional and practical advice. We, all staff in SFP adhere to the Code of Ethics Conduct as outlined below.

**Professional Integrity and Accountability: ** We always serve our clients with integrity and respect. No fabricated or altered data that deliberately affects the results will be used. We promise to be fully qualified with accepted work and be honest to clients if any limitation of competence exists. Appropriate and relevant statistical methods are always used to produce reliable results. All results and conclusions should be clearly explained in detail with statistical data and valid analytic approaches. 

**Responsibilities to clients: ** We treat all clients’ information confidential and responsible. No client’s information should be disclosed or discussed in public without permissions of the clients. We will protect any sensitive data or results from unauthorized people and will not take advantage of them to make profit. 

**Responsibilities to other statisticians: ** We always safeguard the reputation of statistical practice and try to enhance our professionalism. We support and help strengthen the work of other fellow statisticians through valuable peer view. No criticism is allowed to cast on individuals rather than on problems. Even in debates, we should be respectful of others.



